\AuthorRefNames [CHGN]
\begin{references}

\bibitem{Bea}  \name{A.\ Beardon}, {\it The Geometry of Discrete
Groups}, {\it Grad.\  Texts in Math\/}.\  {\bf 91}, Springer-Verlag, 
New York, 1983.
 
 \bibitem{Ber}  \name{J.\ Berge}, {\it Heegaard}, A program to understand Heegaard splittings of
$3$-manifolds, available free from the author.
 
 \bibitem{CHW} \name{P.\ J.\  Callahan, M.\ V.\  Hildebrand}, and \name{J.\ R.\  
Weeks}, A census of cusped hyperbolic $3$-manifolds, {\it Math.\ Comp.\/} 
{\bf 68} (1999), 321--332.
 
 \bibitem{CGHN}  \name{D.\ Coulson, O.\ A.\  Goodman, C.\  Hodgson}, 
and \name{W.\ D.\  Neumann}, Computing arithmetic invariants of $3$-manifolds, 
{\it Experiment.\ Math\/}.\ {\bf 9} (2000), 127--152.
 
 \bibitem{CS} \name{M.\ Culler} and \name{P.\ Shalen}, Varieties of group
representations and splittings of\break $3$-manifolds, {\it Ann.\  of Math\/}.\  
{\bf 117} (1983), 109--146.
 
 \bibitem{F}  \name{W.\ Fenchel},  {\it Elementary Geometry in 
Hyperbolic Space\/}, {\it de Gruyter Studies in Math\/}.\ {\bf 11}, 
Walter de Gruyter \& Co., Berlin (1989).
 
 \bibitem{FS} \name{L.\ H.\ de Figueiredo} and \name{J.\ Stolfi}, Self-validated
numerical methods and applications, Brazilian Math.\  Colloq.\  Monograph, IMPA, Rio de Janeiro, Brazil (1997).
 
 \bibitem{G} \name{D.\ Gabai}, On the geometric and topological rigidity
of hyperbolic $3$-manifolds, {\it J.\ Amer.\ Math.\ Soc\/}.\  
{\bf 10} (1997), 37--74.
 
\bibitem{GM1} \name{F.\ Gehring} and \name{G.\ Martin}, Inequalities for
M\"obius transformations and discrete groups, {\it J.\ Reine Agnew.\
Math\/}.\ {\bf 418} (1991), 31--76.
 
\bibitem{GM2} \bibline, Precisely invariant collars 
and the volume of hyperbolic $3$-folds, {\it J.\ Differential Geom\/}.\ 
{\bf 49}(1998), 411--435.
 
 \bibitem{HW} \name{C.\ Hodgson} and \name{J.\ Weeks}, Symmetries, isometries
and length spectra of closed hyperbolic $3$-manifolds, 
{\it Experiment.\  Math.\/} {\bf 3} (1994), 261--274.
 
 \bibitem{IEEE} IEEE Standard for binary floating-point arithmetic
(ANSI/IEEE Std 754-1985) published by the Institute of 
Electrical and Electronics Engineers, Inc., New York, NY, 1985.
 
 \bibitem{JR} \name{K.\ Jones} and \name{A.\ Reid}, Vol3 and other exceptional 
hyperbolic $3$-manifolds,  {\it Proc.\ A.M.S.} {\bf 129}
(2001), 2175--2185.
 
 \bibitem{K1}  \name{W.\ Kahan}, Interval arithmetic options in the 
proposed IEEE floating point arithmetic standard (Karl L.\ E.\ Nickel,
ed.), in  
{\it Interval Mathematics\/}, 99--128 (1980). 
 
 \bibitem{K2}  \bibline, Beastly numbers, preprint.
 
\bibitem{La}  \name{O.\ Lanford}, Computer-assisted proofs in analysis, 
in {\it Proc.\  ICM} (Berkeley, CA)  {\bf 2} (1986), 1385--1394.
 
 \bibitem{Math} {\it Mathematica} 2.0, {\it A System for Doing Mathematics by Computer}, 
Wolfram Research Inc., Champaign,  Ill.
 
 \bibitem{M1} \name{G. R. Meyerhoff}, A lower bound for the volume of 
hyperbolic 
$3$-manifolds, {\it Canad.\ J.\ Math\/}.\ {\bf 39} (1987), 1038--1056.
 
 \bibitem{M2} \bibline,  Sphere-packing and volume in 
hyperbolic $3$-space, {\it Comm.\ Math.\ Helv\/}.\  {\bf 61} (1986),
271--278.
 
 \bibitem{Mo} \name{G.\ D.\ Mostow}, Quasi-conformal mappings in $n$-space 
and the rigidity of hyperbolic space forms, {\it IHES Publ.\ Math\/}.\ 
{\bf 34} (1968), 53--104.
 
 \bibitem{P} \name{A.\ Przeworski}, Tubes in hyperbolic $3$-manifolds, {\it Topology Appl.} {\bf
148/2-3} (2003), 103--122.
 
 \bibitem{R} \name{R.\ Riley}, Applications of a computer implementation
of Poincar\'e's theorem on fundamental polyhedra,
{\it Math.\ Comp\/}.\  {\bf 40} (1983) 607--632.
 
 \bibitem{T} \name{N.\ Thurston},  Finding killerwords, in preparation. 
 


\bibitem{Wa} \name{F.\ Waldhausen}, On irreducible $3$-manifolds which
are sufficiently large, {\it Ann.\ of Math\/}.\ {\bf 87} (1968), 56--88.
 
 \bibitem{W}  \name{J. Weeks}, {\it SnapPea}, available free from the author.

\vglue9pt

\centerline{\elevenpoint\it Added in proof} 

\vglue5pt

\bibitem{A} I. Agol, Volume change under drilling, {\it Geom.\ Topol.\/} {\bf 6} (2002), 905--916.

\bibitem{P2} A. Przeworski, A universal upper bound on density of tube packings in hyperbolic space,
preprint.

\end{references}

 



