\def\bibline{\hbox to30pt{\hrulefill}}

\begin{thebibliography}{GHMTY}

%\bibitem [A]{A} I. Agol, Volume change under drilling, {\textit Geom.\ Topol.\/} {\textbf 6} (2002), 905--916.

\bibitem[ACS]{ACS} I. Agol, M. Culler, and P. Shalen, Dehn surgery, homology and hyperbolic volume,
\textit {Algebr. Geom. Topol.} \textbf {6} (2006), 2297--2312.

\bibitem [ADST]{ADST} I. Agol, P. Storm \& W. Thurston (appendix by N. Dunfield), Lower bounds on volumes of hyperbolic Haken 3-manifolds, \textit {J. AMS}, \textbf {20} (2007), 1053-1077. 

\bibitem[Bea]{Bea}  A. Beardon, \textit {The Geometry of Discrete Groups}, Graduate Texts in mathematics \textbf {91}, Springer-Verlag, 
%New York, 1983.


%\noindent\hang [Be]  J. Berge, Heegaard, a program to understand Heegaard splittings of
%3-manifolds, available free from the author.

\bibitem [CLLMR]{CLLMR} A. Champanerkar, J.  Lewis, M. Lipyanskiy, S. Meltzer, (appendix by A. Reid)
Exceptional regions and associated exceptional hyperbolic 3-manifolds, \textit {Experiment. Math.} \textbf {16} (2007), no. 1, 107 - 118.

\bibitem [F]{F}  W. Fenchel,  Elementary Geometry in Hyperbolic Space, \textit {de Gruyter Studies Math.}, \textbf {11}, de Gruyter, Berlin (1989).


\bibitem [FS]{FS} L.\ H.\ de Figueiredo and J.\ Stolfi, Self-validated
numerical methods and applications, \textit {Brazilian Math.\  Colloq.\  Monograph}, IMPA, Rio de Janeiro, Brazil (1997).

 
  
 

\bibitem [G]{G} D. Gabai, On the Geometric and Topological Rigidity of Hyperbolic 3-Manifolds,  to appear, \textit {J. Amer. Math. Soc.\/}  \textbf {10} (1997), no. 1, 37?74.

\bibitem [G1]{G1} D. Gabai, The Smale conjecture for hyperbolic 3-manifolds:  $\mathrm {Isom}(M^3) \cong \mathrm{Diff} (M^3)$, \textit {J. Diff. Geom.}  \textbf {58} (2001), 113-149.


\bibitem [GMM]{GMM} D. Gabai, R. Meyerhoff \& P. Milley, Minimum volume cusped hyperbolic 3-manifolds, \textit {J. AMS} \textbf {22} (2009), 1157-1215.

\bibitem [GMT]{GMT} D.Gabai, R. Meyerhoff \& N. Thurston, Homotopy Hyperbolic 3-Manifolds are Hyperbolic, \textit {Annals of Math.} (2) \textbf {157} (2003), 335--431.

\bibitem[GT]{GT} D. Gabai \& M. Trnkova, Exceptional hyperbolic 3-manifolds, \textit {Comment. Math. Helv}, \textbf {90} (2015), 703-730.

%\noindent \hang [GM1] F. Gehring and G. Martin, Inequalities of M\"obius Transformations and Discrete Groups, {\it j. Reine Agnew. math.\/} {\bf 418} (1991) 31 - 76.
%\medskip
%\noindent \hang [GM2] F. Gehring and G. Martin, Precisely invariant collars and the volume of hyperbolic 3-folds, preprint.
%\medskip


\bibitem [GHMTY]{GHMTY} D. Gabai, R. Haraway, R. Meyerhoff, N. Thurston, A. Yarmola, Hyperbolic 3-manifolds of low cusp volume, preprint.

\bibitem [GMTY]{GMTY} D. Gabai,  R. Meyerhoff, N. Thurston, A. Yarmola, Enumerating Kleinian Groups, to appear, \textit {Proceedings of  Computational Aspects of Discrete Subgroups of Lie Groups}, AMS Contemporary Math. Series.

\bibitem [IEEE]{IEEE} IEEE Standard for binary floating-point arithmetic
(ANSI/IEEE Std 754-1985) published by the Institute of 
Electrical and Electronics Engineers, Inc., New York, NY, 1985.
 
 
% \noindent\hang [HW1] C. Hodgson and J. Weeks,  personal communication.
% 
 \bibitem [JR]{JR} K.\ Jones and A.\ Reid, Vol3 and other exceptional 
hyperbolic $3$-manifolds,  \textit {Proc.\ A.M.S.} \textbf {129}
(2001), 2175--2185.
 
 \bibitem [K1]{K1}  W.\ Kahan, Interval arithmetic options in the 
proposed IEEE floating point arithmetic standard (Karl L.\ E.\ Nickel,
ed.), in  
\textit {Interval Mathematics\/}, 99--128 (1980). 
 
% \bibitem [K2]{K2}  W.\ Kahan, Beastly numbers, preprint 1996.


\bibitem[L]{L} M. Lipyanskiy. A Computer-Assisted Application
of Poincar\'e's Fundamental Polyhedron Theorem.
Preprint, 2002.

\bibitem[LM]{LM} M. Lackenby \& R. Meyerhoff, The maximal number of exceptional Dehn surgeries, \textit {Invent. Math.} \textbf {191} (2013), 341--382.

%\noindent\hang [Math] Mathematica 2.0, A system for doing mathematics by computer, 
%Wolfram Research Inc., Champaign Il.

\bibitem[M1]{M1}G. R. Meyerhoff, A Lower Bound for the Volume of Hyperbolic 
3-Manifolds, \textit {Canadian J. Math.} \textbf {39} (1987) 1038-1056.

%\noindent\hang [M2] G. R. Meyerhoff,  Sphere-Packing and Volume in Hyperbolic 3-Space, {\it Comment. Math. Helvetici\/} {\bf 61} (1986) 271-278.
%\medskip

\bibitem[Mi]{Mi} P. Milley, Minimum volume hyperbolic 3-manifolds, \textit {J. Top.} \textbf {2}(2009), 181-192.

%\noindent\hang [Mo] G. D. Mostow, Quasi-Conformal Mappings in n-Space and the Rigidity of Hyperbolic Space Forms, {\it IHES Publ. Math. \/} {\bf 34} (1968) 53-104.
%\medskip
\bibitem[Pe1]{Pe1} G. Perelman, Ricci flow with surgery on three-manifolds, arXiv:math.DG/0303109.\bibitem[Pe2]{Pe2} G. Perelman, The entropy formula for the Ricci flow and its geometric applications,arXiv:math.DG/0211159.

% \bibitem [P2]{P2} A. Przeworski, A universal upper bound on density of tube packings in hyperbolic space,
% {\textit J. Differential Geom.} {\textbf 72} (2006), no. 1, 113?127.
% 
% \noindent \hang [Wa] F. Waldhausen, On Irreducible 3-Manifolds which Are Sufficiently Large, {\it Annals of Math. \/} {\bf 87} (1968) 56-88.
%\medskip


%\bibitem [W]{W}  J. Weeks, SNAPPEA program.  
\end{thebibliography}

 
